cleaning the pheno table

### 2020-11-10 scripts/37_clean_pheno_refreshed.Rmd
update

```{r}
# write out the asv seq and cts to corresponding files
library(seqinr)
cts_fns <- list(msk = msk_cts,
                italy = italy_cts,
                upenn = upenn_cts)

@@

    # the cts for the asv
    .x %>% 
      write_tsv(str_glue('../data/picrust/input/{.y}_cts.tsv'))

@@

    write.fasta(sequences = as.list(seqs$asv_sequence),
              names = seqs$key,
              file.out = str_glue('../data/picrust/input/{.y}_seqs.tsv'))
  })
```

>>>>

## write out the asv counts and seqs for picrust2

```{r}
# write out the asv seq and cts to corresponding files
# ITS NOT FILTERED BUT TO ALIGN WITH THE SNAKEFILE FILE NAMES
library(seqinr)
cts_fns <- list(msk = msk_cts,
                italy = italy_cts,
                upenn = upenn_cts)

@@

    # the cts for the asv
    .x %>% 
      write_tsv(str_glue('../data/picrust/input/filtered_counts_{.y}.tsv'))

@@

    write.fasta(sequences = as.list(seqs$asv_sequence),
              names = seqs$asv_key,
              file.out = str_glue('../data/picrust/input/filtered_counts_asv_sequence_{.y}.tsv'))
  })
```

## load the downloaded results from picrust2

```{r}
src <- tibble(
  id = c('upenn','msk','italy'),
  local_dir = str_glue('/Users/daia1/projects/CART_microbiome/data/picrust/output/{id}'),
  pw_file = str_glue('daia1@lilac.mskcc.org:~/my_workdir/other_pipeline/picrust2/output/picrust2_out_{id}/pathways_out/path_abun_unstrat.tsv.gz'),
  dl_pw = str_glue('scp {pw_file} {local_dir}'),
  unzip_pw = str_glue('gunzip /Users/daia1/projects/CART_microbiome/data/picrust/output/{id}/path_abun_unstrat.tsv.gz')
)
# create local dir first
src %>% 
  mutate(mkdir_cmd = str_glue('mkdir {local_dir}')) %>% 
  select(mkdir_cmd) %>% 
  write_tsv('../data/mkdir.sh', col_names = F)

src %>% 
  select(dl_pw,  unzip_pw) %>% 
  gather('fn', 'cmd') %>% 
  select(cmd) %>% 
  write_tsv('../data/downloding_pw.sh', col_names = F)
```

## comparison together and individually

Lump italy and msk together since there are only three baseline samples from italy, and also italy and msk observe CR at day 30.

```{r}
pw_msk <- read_tsv('../data/picrust/output/msk/path_abun_unstrat.tsv')
pw_italy <- read_tsv('../data/picrust/output/italy/path_abun_unstrat.tsv')
pw_upenn <- read_tsv('../data/picrust/output/upenn/path_abun_unstrat.tsv')

# msk and italy
pheno_mitaly <- bind_rows(
  msk %>% 
    select(sampleid, CR, Toxicity),
  italy %>% 
    select(sampleid = oligoid,
           CR,
           Toxicity = Tox)
)

pw_mitaly <- pw_msk %>% 
  inner_join(pw_italy)

pw_mitaly_or <- bind_cols(pw_mitaly[,1],pw_mitaly[,pheno_mitaly$sampleid])

all.equal(colnames(pw_mitaly_or)[2:ncol(pw_mitaly_or)], pheno_mitaly$sampleid)

# upenn
pheno_upenn <- upenn %>% 
  mutate(sampleid = as.character(sampleid)) %>% 
  select(sampleid, CR, Toxicity = Tox)

pw_upenn_or <- bind_cols(pw_upenn[,1],pw_upenn[,pheno_upenn$sampleid])

all.equal(colnames(pw_upenn_or)[2:ncol(pw_upenn_or)], pheno_upenn$sampleid)

# all of them
pheno_all <- bind_rows(
  pheno_mitaly,
  pheno_upenn
)

pw_all <- pw_mitaly %>% 
  inner_join(pw_upenn)

pw_all_or <- bind_cols(pw_all[,1],pw_all[,pheno_all$sampleid])

all.equal(colnames(pw_all_or)[2:ncol(pw_all_or)], pheno_all$sampleid)
```
```{r}
pheno <- list(mitaly = pheno_mitaly,
              upenn = pheno_upenn,
              all = pheno_all)

pheno %>% 
  imap(~ select(.data = .x, CR) %>% 
            t %>% 
            write.table(str_glue('../data/lefse/{.y}__pheno_CR.txt'), sep = '\t', quote = F, row.names = T, col.names = F))

pheno %>% 
  imap(~ select(.data = .x, Toxicity) %>% 
            t %>% 
            write.table(str_glue('../data/lefse/{.y}__pheno_Toxicity.txt'), sep = '\t', quote = F, row.names = T, col.names = F))

cts_or <- list(
              mitaly = pw_mitaly_or,
              upenn = pw_upenn_or,
              all = pw_all_or
)

cts_or %>% 
  imap(~ .x %>% write_tsv(str_glue('../data/lefse/{.y}_cts.tsv')))

```{r}
# a script to bind those things and produce the correct format for the galaxy portal
cts <- list.files('../data/lefse/', pattern = 'cts')
pheno <- list.files('../data/lefse/', pattern = 'pheno')
combi <- expand.grid(cts_ = cts,pheno_ = pheno) %>% 
  mutate(cts_id = str_replace(cts_, '_cts.tsv','')) %>% 
  mutate(pheno_id = str_replace(pheno_, '__.+$','')) %>% 
  filter(cts_id == pheno_id) %>% 
  mutate(grp = if_else(str_detect(pheno_, 'CR'), 'CR', 'tox')) %>% 
  mutate(cmd = str_glue('cat {pheno_} {cts_} > {pheno_id}_{grp}.tsv'))

combi %>% select(cmd) %>% 
  write_tsv('../data/lefse/cat_table.sh', col_names = F)
